---
title: "Where your vote counts"
date: "15/04/2022"
output: 
  html_document: 
    fig_width: 8
    fig_height: 6
    includes:
       in_header: analytics.html
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



```{r dataproc, include=FALSE}
library(eechidna)
library(tidyverse)
library(purrr)
library(purrrlyr)
library(scales)
library(spatial)
library(geojsonio)
library(spdplyr)
library(rmapshaper)
library(viridis)
library(sf)
library(leaflet)
library(htmltools)


# data(abs2016)
# data(fp16)
# data(tpp16)
# data(tcp16)
# fp=fp16
# abs=abs2016
# tpp=tpp16
# tcp=tcp16

data(abs2016)
data(fp19)
data(tpp19)
data(tcp19)
abs=abs2016
fp=fp19
tpp=tpp19
tcp=tcp19

# nat_map = readRDS(file = "nat_map.Rdata")
nat_sff = readRDS(file = "nat_sff.Rdata")
nat_data = readRDS(file = "nat_data.Rdata")

# # inspect the data frame
# glimpse(fp)
# # show the first few rows
# head(fp) %>% kable

# who_most_votes_prop <- fp %>% 
#     filter(CandidateID != 999) %>% #exclude informal votes
#     mutate(candidate_full_name = paste0(GivenNm, " ", Surname, " (", CandidateID, ")")) %>% 
#     group_by(candidate_full_name) %>% 
#     dplyr::summarise(total_votes_for_candidate = sum(Percent, rm.na = TRUE)) %>% 
#     arrange(desc(total_votes_for_candidate)) %>% 
#     dplyr::rename(percent_votes_for_candidate = total_votes_for_candidate)

# partycolour_map = c(
#     "AUSTRALIAN LABOR PARTY" = "#FF0033",
#     "LIB/NATS COALITION" = "#0066CC",
#     "INDEPENDENT" = "#000000",
#     "KATTER'S AUSTRALIAN PARTY (KAP)" = "#CC3300",
#     "PALMER's UNITED AUSTRALIA PARTY" = "#FFFF00",
#     "CENTRE ALLIANCE" = "#883388",
#     "THE GREENS" = "#009900"
# )

cart.winners <- fp %>% 
    filter(Elected == "Y") %>% 
    select(DivisionNm, PartyNm,) %>% 
    mutate(PartyNm = ifelse(PartyNm %in% c("LIBERAL PARTY", "NATIONAL PARTY"), "LIB/NATS COALITION", PartyNm)) %>% 
    merge(nat_data, by.x="DivisionNm", by.y="elect_div")

# Plot it inside R
# ggplot(data=nat_map) +
#     geom_polygon(aes(x=long, y=lat, group=group),
#                  fill="grey90", colour="white") +
#     geom_point(data=cart.winners, aes(x=x, y=y, colour=PartyNm), size=1.5, alpha=0.8) +
#     scale_colour_manual(name="Political Party", values=partycolour_map) +
#     theme_map() + coord_equal() + theme(legend.position="bottom")

tcpp = left_join(
    tcp %>% 
        distinct(CandidateID, .keep_all=TRUE) %>% #duplicate rows for some reason?
        rename(TCP_Percent=Percent, TCP_Votes=OrdinaryVotes),
    tpp %>% 
        select(DivisionNm, LNP_Votes, LNP_Percent, ALP_Votes, ALP_Percent, TotalVotes, Swing),
    by=c("DivisionNm"))

vp = left_join(
    tcpp,
    fp %>%
        distinct(CandidateID, .keep_all=TRUE) %>% # weird duplicates. why?
        select(CandidateID, OrdinaryVotes, Percent) %>%
        rename(FP_Percent=Percent, FP_OrdinaryVotes=OrdinaryVotes), 
    by=c("CandidateID")
)

vp.winners = vp %>% filter(Elected=="Y") %>%
    mutate(Margin=ceiling(TCP_Votes-TotalVotes/2)) %>%
    # mutate(QMargin=1-quantile.t(Margin)) %>%
    mutate(Margin_Prop=Margin/TotalVotes) # %>%
    # mutate(QFP_Percent=1-quantile.t(FP_Percent))

winners_sf = left_join(
    nat_sff %>%  dplyr::rename(DivisionNm = elect_div),
    vp.winners,
    by=c("DivisionNm"),
    copy=TRUE)
```

```{r drawmap, echo=FALSE}
rawleafletmap <- leaflet(data=winners_sf) %>%
    addProviderTiles(
        "CartoDB.Positron",
        options = tileOptions(
            minZoom=4, 
            maxZoom=12
        )
    )

pal <- colorNumeric(
    "magma",
    range(winners_sf$Margin),
    reverse=TRUE
)

popupHtml = apply(winners_sf, 1, function(r){
    s = as.character(withTags({
        div(
            p(
                h3(str_to_title(paste("Electoral Division:", r$DivisionNm)))
            ),
            p(
                str_to_title(paste(r$GivenNm, r$Surname, paste(sep="", '(', r$PartyNm, ')')))
            ),
            dl(
                dt('Winnning Margin'), dd(r$Margin ),
                dt('First preference votes'), dd(r$FP_OrdinaryVotes),
                dt('Total Votes'), dd(r$TotalVotes),
            )
        )
    }))
    # s[[1]]
})


leafletmap <- rawleafletmap %>%
    addPolygons(
        fillColor = ~pal(Margin),
        weight = 1,
        opacity = 1.0, fillOpacity = 0.5,
        label = ~as.character(str_to_title(DivisionNm)),
        highlightOptions = highlightOptions(
            color = "white",
            weight = 2,
            bringToFront = TRUE
        ),
        dashArray='3'
    ) %>% addLegend(
        pal = pal,
        values = ~Margin,
        opacity = 0.5,
        title="Winning margin",
        # labFormat = labelFormat(transform = function(x) round(10^x))
    ) %>% addMarkers(
        ~long_c, ~lat_c, 
        popup = popupHtml,
    )  #%>% htmlwidgets::onRender( # see https://rstudio.github.io/leaflet/morefeatures.html#custom-javascript-with-htmlwidgetonrender
    #     "function(el, x) {
    #       var leafletmap = this;
    #       leafletmap.on('baselayerchange',
    #         function (e) {
    #           leafletmap.minimap.changeLayer(L.tileLayer.provider(e.name));
    #         })
    #     }"
    # )
leafletmap
```


Light colours:
: Marginal seats. You vote has a lot of power to influence policy and government.

Dark colours:
: Non marginal seats. Your representative is secure and does not need your support all that much thanks.

Next time you feel like you do not have a say in how the government operates, maybe you could think to yourself “maybe I just don’t live in the right place”?

In Australia, the party that gets to form government is the one `who wins the most seats in the House of Representatives, which comes down to winning the vote in the majority of federal electorates.
But the competitions are much closer for some seats than others.
How important your vote is depends on where that vote is,
In some electorates the margin is very narrow, and the victorious candidate is chosen by no more voters than you can fit in a pub.

Or, to turn that around, if you live in one of those areas where one pub-full of people get to choose who wins, your vote is very important, and you have a representative who really needs your support to keep their job.
Next time you feel like the government is ignoring you, maybe you should consider moving to one of these _marginal seats_ and then see if your representative isn’t substantially more interested in hearing what you have to say.
Either way, if you do want to have a say in how this government is run, it won’t hurt to get a pub full of mates to make the case with you (or a church full, or a book-club full, or a mosque-full…

So what is on this map?
What we have here are the electorates from the 2019 Federal election, colourised by the size of the margin that the winning candidate had,
Assuming that voting patterns are similar in the next election, this can indicate roughly how much your vote matters in the next election too.

## Technical details

If you want to see how it is done, or to make your own version, visit the 
[source code at github](https://github.com/oz-info/ozpowermap/).
You can make one of your own.
Why not build one for your state or territory?
Or one for the Senate instead of the House of Reps?
Or make it look prettier? We are not graphic designers over here. 😜
Or do a more sophisticated analysis! There is a lot more data to play with here and we are only scratching the surface.

## TODO

This is a work in progress.
We could do lots of things to make this better

1. Those map markers get in the way; maybe we could replace them with a mouseover thing
2. perhaps this would be better as an overlay on google maps, so you can search and navigate electorates better
3. Contributor instructions could be clearer
4. … whatever you want! This project is open source and you can change it!

## Credits

A project started by [Dan MacKinlay](https://danmackinlay.name).

This site is built upon work, tutorials, inspiration and/or advice from many people we offer them thanks to their inspiration.

* [Jeremy Forbes](https://github.com/jforbes14/) and the [eechidna contributors](https://github.com/jforbes14/eechidna/graphs/contributors)
* [Ben Raue](https://www.tallyroom.com.au/) 
* [Timothy Lin](https://www.timlrx.com/blog/using-leaflet-in-r-tutorial)
* [Peter Ellis](http://freerangestats.info/)
* [Dr Kevin Bonham](https://kevinbonham.blogspot.com/)

However, all mistakes made here and opinions expressed here are the responsibility of the ozpowermap project contributors, not these folk.
